Seasonality in governing dynamics at submesoscales

### 2016-06-10 writeup/rocha_etal.tex
Makes various small edits + adds appendix

\begin{abstract}
    Two new high-resolution numerical simulations with embedded tides suggest a
    strong modulation in the governing near-surface dynamics at submesoscales
    (roughly 10-100 km). Consistent with recent studies, deep late Winter mixed layers
    are prone to baroclinic instabilities and submesoscale turbulence
    prevails in late Winter/early Spring. Inertia-gravity waves, on the other hand, are enhanced near
    the surface when the upper ocean in strongly stratified. As a consequence of
    reduction of submesoscale turbulence and enhancement of intertia-gravity waves,
    the latter strongly dominate the submesoscale surface kinetic energy and
    sea-surface height variance in late Summer/early Fall.
\end{abstract}

@@

Using the output of two new $1/24^\circ$ and $1/48^\circ$ global
numerical simulations with embedded tides, we show that inertia-gravity waves undergo
a strong seasonality near the surface in the Kuroshio Extension region.
The seasonal cycle of inertia-gravity waves is out-of-phase
with the seasonal cycle of submesoscale turbulence. Consistent with previous studies,
deep late Winter mixed layers are prone to
shallow baroclinic instabilities that are roughly in geostrophic balance
and flux energy upscale \citep{sasaki_etal2014,callies_etal2016},
driving a mild seasonal modulation of the mesoscales \citep{sasaki_etal2014,qiu_etal2014}.

@@

\section{The LLC numerical simulations}
We use the output of two latitude-longitude polar cap (LLC)
MITgcm realstic numerical simulations. The outputs
 analyzed here, LLC2160 and LLC4320,
are the 1/24$^\circ$ and 1/48$^\circ$  --- 90-vertical-levels simulations from
a hierarchy of global
forward numerical solutions on a LLC grid   \citep{forget_etal2015} that were
spun up from a ECCO2 adjoint-method state estimate. Both simulations were forced by
tides and 6-hourly surface fluxes. The LLC2160
output spans two years from March 2011 to April 2013; the LLC4320 was spun up from
the LLC2160 simulation, spanning one year from September 2011 to September 2012.
The LLC4320 simulation is an extension of the then 5-month long output used by
\citet{rocha_etal2016}.

@@

To study seasonal variations in the upper-ocean dynamics, we focus on the northwest
Pacific, in the vicinities of the Kuroshio
Extension, where previous studies have suggested strong submesoscale seasonality.
We analyze a sub-domain of the LLC4320 and LLC2160 simulations of about 2000 km$^2$
spanning 155-175$^\circ$E; 25-40$^\circ$N (figure \ref{fig1}a). The stratification
in this mesoscale-rich subtropical region undergoes a strong seasonal cycle: wintertime
enhance small-scale turbulence de-stratify the upper ocean, yielding mixed layers
as deep as 300m. In late Spring/early Summer enhanced solar radiation and mixed-layer
instabilities re-stratify the upper ocean, yielding mixed layers as shallow as 40m.
The stratification and its seasonality is well-captured by both LLC simulations:
the skillful comparisons of model stratification to Argo climatology lends confidence
to the present analysis (supplemental material).

@@

Figures \ref{fig1}b-c show snapshots of vertical vorticity $\zeta$ in early Spring
(April 15) and Fall (October 15) in the LLC4320 (1/48$^\circ$) simulation.
The model solutions depict seasonality in vorticity: large values of
fine-grained vertical vorticity are observed in early Spring with maximum
values as large as $4f$, where $f$ is
the local planetary vorticity, and root-mean-square (RMS) of about 0.4. In early
Fall, the situation is the opposite: the vertical vorticity is relatively coarse-grained,
and its local maximum and RMS are no larger than $0.5f$ and $0.2f$, respectively.
Indeed, the seasonal cycle of vorticity and rate of strain is very strong (figure~\ref{fig2}):
in both simulations, the RMS vorticity and strain are about twice as large in
late Winter/early Spring
than in late Summer/early Fall. Because the wintertime vorticity and strain rate
are dominated by the smallest scales in the flow, increasing the resolution from
1/24$^\circ$ to 1/48$^\circ$ significantly increases the wintertime RMS vorticity
and strain by about 40$\%$.

@@

The bulk of vertical vorticity and lateral strain rate is associated with subinertial flows
($T_f = 2\pi/f_0\approx 23.5$h, where $f_0$ is the inertial frequency at
the mean latitude): daily-averaging the velocity fields reduces the RMS
vorticity by less than 40$\%$ and the RMS strain by less than 10$\%$; the seasonal
cycle remains strong (see
dashed lines in figures \ref{fig2}a-b). Most of this seasonal cycle is associated
with submesoscale flows: smoothing the velocity fields
with a Hanning filter with cut-off scale of 100 km dramatically reduces the RMS
vorticity and strain. The reduction in variance is about 80$\%$ in winter,
yielding RMS vorticity and strain roughly consistent the ones diagnosed from
AVISO gridded geostrophic velocities (compare blue lines with black lines in figures \ref{fig2}a-b).
The overall picture that emerges is consistent with recent studies: shallow baroclinic
instabilities energize the submesoscales in late Winter, drawing from the available
potential energy stored in deep mixed layers \citep{sasaki_etal2014,callies_etal2015,callies_etal2016}.

The seasonal cycle of the horizontal divergence, however, showcases the complexity
of the upper-ocean annual variability. If submesoscale eddies and fronts dominated
the near-surface variability all year, then the seasonal cycle of horizontal divergence,
vertical vorticity, and lateral strain rate would be in phase \citep[e.g.,][]{sasaki_etal2014}.
While there is a clear wintertime peak in divergence of daily-averaged velocity
(see dashed lines in figure \ref{fig2}a; RMS divergence $\sim$0.22 f in the 1/48$^\circ$
simulation), the hourly fields show
a stronger enhancement of
lateral divergence in late Summer/early Fall (RMS divergence $\sim$0.22 f the 1/48$^\circ$
simulation). Because the 1/48$^\circ$ simulation better resolve
smaller-scale submesoscale flows, the
winter peak in divergence is nearly as strong the summertime's. Of course, part of
these submesoscale fronts and eddies evolve relatively fast and there is no clear
temporal and spatial scale separation between those motions and inertia-gravity waves:
daily-averaging the velocity fields efficiently suppresses the summertime horizontally
divergent flows, but also reduces the wintertime lateral divergence by about 50$\%$.
As for the vorticity and strain rate, most of the lateral divergence is associated
with submesoscale flows: smoothing the the velocity fields with a 100km-cut-off
suppresses more than 80$\%$ of the RMS divergence.

\section{Seasonality in governing submesoscale dynamics}
The overall results of figure \ref{fig2}c strongly suggest that the
summertime submesoscale surface variability stems from different dynamics than
then wintertime variability. To better understand these differences, we calculate
vorticity-divergence and vorticity-strain joint probability distributions (jPDF).

The April vorticity-strain jPDF have a shape characteristic of submesoscale turbulence
 (see figures \ref{fig3}f-g). The alignment of vorticity and strain $\alpha \sim \zeta$
 with strong positive skewness are fingerprints of submesoscale fronts
 \citep{shcherbina_etal2013,mcwilliams2016}. The shape of the vorticity-strain
 jPDF is similar for hourly, daily-averaged, and residual velocities, although
 the vorticity skewness reduces 1.4 to 1.13 from hourly to daily-averaged.
 The April vorticity-divergence jPDFs, however, have a egg-like shape that which again highlight
 the strong skewness in vorticity. Thus wintertime submesoscale surface velocity
 is strongly dominated by submesoscale turbulence --- this is true even for
 superinertial submesoscale currents.

The October vorticity-strain jPDF show much weaker skewness (the vorticity skewness
is 0.68 and 0.67 for hourly and daily-averaged velocities). The shape of the
 vorticity-strain jPDF  appears to be a combination of two half-ellipses centered
 about $\zeta=0$, one with a 45$^\circ$ slope (characteristic of submesocale fronts
 that persist in Summer) and one with a very steep slope. The vorticity-divergence
 jPDF is significantly different than the April one. In particular, the jPDF
 for daily-averaged currents is still egg-shaped but hourly and residual are
 ellipses about $\zeta=0$, likely associated with inertia-gravity waves.

@@

To better understand the projection of these flows onto different horizontal
scales, we calculate wavenumber spectra of kinetic energy and sea-surface height
(SSH) variance.  Before calculating the spectra, time-mean and and spatial linear
trend were removed, and the ensuing fields were multiplied by a Hanning spectral ``window''.
We discuss only spectra for the 1/48$^\circ$, which simply extends the 1/24$^\circ$
towards smaller scales (supplemental material).

Figure \ref{fig3}a depicts the horizontal wavenumber spectra of surface KE in
April and October. At scales larger than 20km, the spectra based on hourly
velocity snapshots are nearly indistinguishable within 95$\%$ confidence level
(compare solid lines in figure \ref{fig3}a). Consistent with the results of
\citet{rocha_etal2016} who analyzed a 5-month output of the LLC4320 simulation in Drake Passage,
there is significant high-frequency variabilty at submesoscales. Daily-averaging
the velocity field suppresses spatial variability at scales smaller than about 250
km both in April and October (compare solid lines against dashed lines in figure
\ref{fig3}a). But this suppression is more dramatic in October, when the horizontally
divergent flows peak. At scales
smaller than 100 km, 39$\%$ of the surface KE in April is accounted for by super-inertial
flows as opposed to 79$\%$ in October. The seasonality of subinertial submesoscale flows
is much more dramatic than the seasonal cycle of the total flow. This is consistent
with the results of \citet{sasaki_etal2014}, which are
based on daily-averaged velocity fields of a different model without tidal forcing
(P. Klein, personal communication).

These high-frequency submesoscale flows significantly project on the sea-surface.
In October, when the inertia-gravity waves are stronger near the surface and submesoscale
turbulence is weak, there is a dramatic difference between the spectra
based on hourly and daily-averaged SSH: at scales smaller than 100 km, the spectra
of hourly SSH roughly follows a $-2$ power-law, whereas the spectra of daily-averaged
SSH roughly follows a $-5$ power-law. At scales
smaller than 100 km, 33$\%$ of the SSH variance in April is accounted for by suprainertial
flows as opposed to 83$\%$ in October! The out-of-phase seasonal cycle of submesoscale
turbulence and near-surface inertia-gravity waves conspire to yield very weak seasonality
in the spectra of KE and SSH variance based on hourly field.

\section{Summary and Conclusion}
Our main finding is that
near-surface upper-ocean inertia-gravity waves, undergo
a strong seasonal cycle that is out-of-phase with the seasonal cycle of
the submesoscale turbulence. \cite{dasaro1978} showed that linear internal waves
velocity in the mixed-layer strongly depends on the density jump at the mixed-layer
base. In Summer, the shallow mixed-layer overlays a strong seasonal pycnocline,
and thus the internal waves projection onto the mixed-layer may be stronger
according to \cite{dasaro1978}'s arguments. An
alternative explanation is that the shape baroclinic modes experience significant
seasonal modulation, with significant summertime peak in their surface amplitude.
If the internal wave source has weak seasonal dependence, then a strong
seasonal cycle in the near-surface expression of internal waves is expected.
Internal tide generation is unlikely to vary significantlly seasonally \citep[e.g.,][]{alford2003},
and therefore it is plausible to expect a strong seasonality in the the near-surface
expression of internal tides and
other small-scale internal waves generated through internal tide interactions.
Near-inertial wave generation peaks in Winter, but those waves project on large
horizontal scales \citep[e.g, ][]{qi_etal1995}.

@@

%  this should go in a supplementary material
\clearpage

\appendix

\section{Supplemental material 1: Spectral errors}

\section{Supplemental material 2: On the LLC simulations}

The time-step was 45s for
the LLC2160 simulation and 25s for the LLC4320 simulation. Details of the LLC simulations
are provided in the supplemental material.

 LLC hierarchy setup
and control files are available online
(at \texttt{http://mitgcm.org/viewvc/MITgcm/MITgcm$\_$contrib/
llc$\_$hires}). For the LLC hierarchy, the sea-surface height is calculated
using a linearized equation. While the nominal horizontal resolution  of the
LLC2160 simulation is 1/24$^\circ$,
 horizontal wavenumber spectra suggests an effective resolution of about 20 km.
 (see figure \ref{fig3}).

\subsection{Seasonality in the upper-ocean stratification}

@@

\subsection{Spectra}

\subsection{Seasonality in statistics: probability density functions}
\begin{figure*}[ht]
   \begin{center}
     \includegraphics[width=1.\textwidth]{figs/fig_s2_2.pdf}
  \caption{Time dependence of probability density functions of vertical vorticity (a, b)
  and horizontal divergence (c, d). Monthly averages in April and October are
  also shown (e, f).}
  \label{figS2_2}
  \end{center}
\end{figure*}

\subsection{How well does the model captures high-frequency modes?}
To assess how well the LLC2160 represents the high-frequency variability,
we compare the model near-surface velocity field against two available
mooring data: KEO (the data is available online at \texttt{http://www.pmel.noaa.gov/ocs/KEO})
and KESS mooring $\#$7 (the data is available online at \texttt{http://uskess.whoi.edu}).
For proper comparison, we use model data closest to these two moorings (west of
the region considered in the present study). Focusing on high-frequencies, we
calculate frequency spectral estimates every month. There are about 48 independent
spectral realizations.

@@

\section{Supplemental material 3: Seasonality in pressure modes near-surface amplitude}

>>>>

\begin{abstract}
    Two new high-resolution numerical simulations with embedded tides suggest a
    strong modulation in the governing near-surface dynamics at submesoscales
    (roughly 10-100 km). Consistent with recent studies, deep late Winter mixed layers
    are prone to baroclinic instabilities and submesoscale turbulence
    prevails in late Winter/early Spring. Inertia-gravity waves, on the other hand, are enhanced near
    the surface when the upper ocean in strongly stratified. Thus the summertime
    re-stratification weakens submesoscale turbulence and enhances intertia-gravity
     waves near the surface. In the Kuroshio Extension region,
    inertia-gravity waves strongly dominate the surface kinetic energy and
    sea-surface height variance in late Summer/early Fall at scales as large as 80 km!
\end{abstract}

@@

Using the output of two new $1/24^\circ$ and $1/48^\circ$ global
numerical simulations with embedded tides, we show that inertia-gravity waves undergo
a strong seasonality near the surface in the Kuroshio Extension region.
Interestingly, the seasonal cycle of inertia-gravity waves is out-of-phase
with the seasonal cycle of submesoscale turbulence. Consistent with previous studies,
deep late Winter mixed layers are prone to
shallow baroclinic instabilities that are roughly in geostrophic balance
and flux energy upscale \citep{sasaki_etal2014,callies_etal2016},
driving a mild seasonal modulation of the mesoscales \citep{sasaki_etal2014,qiu_etal2014}.

@@

\section{The LLC numerical simulations}
We use the output of two latitude-longitude polar cap (LLC)
MITgcm realstic numerical simulations. The outputs
 analyzed here, LLC2160 and LLC4320,
are the 1/24$^\circ$ and 1/48$^\circ$  --- 90-vertical-levels simulations from
a hierarchy of global
forward Massachusetts Institute of Technlogy general circulation model (MITgcm)
numerical solutions on a LLC grid   \citep{forget_etal2015} that were
spun up from a ECCO2 adjoint-method state estimate. Both simulations were forced by
tides and 6-hourly surface fluxes. The LLC2160
output spans two years from March 2011 to April 2013; the LLC4320 was spun up from
the LLC2160 simulation, spanning one year from September 2011 to September 2012.
The LLC4320 simulation is an extension of the then 5-month long output used by
\citet{rocha_etal2016}.

@@

To study seasonal variations in the upper-ocean dynamics, we focus on the northwest
Pacific, in the vicinities of the Kuroshio
Extension, where previous studies have suggested strong submesoscale seasonality
\citep{sasaki_etal2014,qiu_etal2014}.
We analyze a sub-domain of the LLC4320 and LLC2160 simulations of about 2000 km$^2$
spanning 155-175$^\circ$E; 25-40$^\circ$N (figure \ref{fig1}a). The stratification
in this mesoscale-rich subtropical region undergoes a strong seasonal cycle: wintertime
enhance small-scale turbulence de-stratify the upper ocean, yielding mixed layers
as deep as 300m. In late Spring/early Summer enhanced solar radiation and mixed-layer
instabilities re-stratify the upper ocean, yielding mixed layers as shallow as 40m.
The stratification and its seasonality is well-captured by both LLC simulations:
the skillful comparisons of model stratification to Argo climatology lends confidence
to the present analysis (supplemental material).

@@

Figures \ref{fig1}b-c show snapshots of vertical vorticity $\zeta$ in early Spring
(April 15) and Fall (October 15) in the LLC4320 (1/48$^\circ$) simulation.
The model solutions depict seasonality in vorticity: large values of
fine-grained vertical vorticity are observed in early Spring with maximum
values as large as $4f$, where $f$ is
the local planetary vorticity, and root-mean-square (RMS) of about 0.4 f. In early
Fall, the situation is the opposite: the vertical vorticity is relatively coarse-grained,
and its local maximum and RMS are no larger than $0.5f$ and $0.2f$, respectively.
Indeed, the seasonal cycle of vorticity and rate of strain is very strong (figure~\ref{fig2}):
in both simulations, the RMS vorticity and strain are about twice as large in
late Winter/early Spring
than in late Summer/early Fall. Because the wintertime vorticity and strain rate
are dominated by the smallest scales in the flow, increasing the resolution from
1/24$^\circ$ to 1/48$^\circ$ significantly increases the wintertime RMS vorticity
and strain by about 40$\%$.

@@

The bulk of vertical vorticity and lateral strain rate is associated with subinertial flows
($T_f = 2\pi/f_0\approx 23.5$h, where $f_0$ is the inertial frequency at
the mean latitude): daily-averaging the velocity fields reduces the RMS
vorticity by less than 40$\%$ and the RMS strain by less than 10$\%$; the seasonal
cycle remains strong (see
dashed lines in figures \ref{fig2}a-b). Most of this seasonal cycle is associated
with submesoscale flows: smoothing the velocity fields
with a Hanning filter with cut-off scale of 100 km dramatically reduces the RMS
vorticity and strain. The reduction in variance is about 80$\%$ in winter,
yielding RMS vorticity and strain roughly consistent with the diagnostics from
AVISO gridded geostrophic velocities (compare blue lines with black lines in figures \ref{fig2}a-b).
The overall picture that emerges is consistent with recent studies: shallow baroclinic
instabilities energize the submesoscales in late Winter, drawing from the available
potential energy stored in deep mixed layers \citep{sasaki_etal2014,callies_etal2015,callies_etal2016}.

The seasonal cycle of the horizontal divergence, however, showcases the complexity
of the upper-ocean annual variability. If submesoscale eddies and fronts dominated
the near-surface variability all year, then the seasonal cycle of horizontal divergence,
vertical vorticity, and lateral strain rate would be in phase \citep[e.g.,][]{sasaki_etal2014}.
While there is a clear wintertime peak in divergence of daily-averaged velocity
(see dashed lines in figure \ref{fig2}a; RMS divergence $\sim$0.22 f in the 1/48$^\circ$
simulation), the hourly fields show
a stronger enhancement of
lateral divergence in late Summer/early Fall (RMS divergence $\sim$0.22 f the 1/48$^\circ$
simulation). Because the 1/48$^\circ$ simulation better resolve
smaller-scale submesoscale flows, the
winter peak in divergence is nearly as strong the summertime's. Of course, part of
these submesoscale fronts and eddies evolve relatively fast and there is no clear
temporal and spatial scale separation between those motions and inertia-gravity waves
\citep{mcwilliams2016}:
daily-averaging the velocity fields efficiently suppresses the summertime horizontally
divergent flows, but also reduces the wintertime lateral divergence by about 50$\%$.
As for the vorticity and strain rate, most of the lateral divergence is associated
with submesoscale flows: smoothing the the velocity fields with a 100km-cut-off
suppresses more than 80$\%$ of the RMS divergence.

\section{Seasonality in governing submesoscale dynamics}
The overall results of figure \ref{fig2}c strongly suggest that the
summertime submesoscale surface variability stems from different dynamics than
then wintertime variability. To better understand these differences, we calculate
joint probability distributions (jPDF) of
vorticity-divergence, vorticity-strain, and vorticity-laplacian of sea-surface height.

The April vorticity-strain jPDF have a shape characteristic of submesoscale turbulence
 (see figures \ref{fig3}f-g). The alignment of vorticity and strain $\alpha \sim \zeta$
 with strong positive skewness are fingerprints of submesoscale fronts
 \citep{shcherbina_etal2013,mcwilliams2016}. The shape of the vorticity-strain
 jPDF is similar for hourly, daily-averaged, and residual velocities, although
 the vorticity skewness reduces 1.4 to 1.13 from hourly to daily-averaged.
 The April vorticity-divergence jPDFs, however, have a egg-like shape that which again highlight
 the strong skewness in vorticity. Thus wintertime submesoscale surface velocity
 is strongly dominated by submesoscale turbulence --- this is true even for
 superinertial submesoscale currents. The hourly velocity fields are significantly
ageostrophic as depicted by the jPDF of vorticity-laplacian of SSH
(figure \ref{fig3}o); it is only the daily-averaged velocity that is approximately
in geostrophic balance (figure \ref{fig3}l). While there is certainly much to be
learned using quasi-geostrophic models to study submesoscale turbulence, claims that
the latter is dominantly in geostrophic balance \cite{sasaki_etal2014,callies_etal2016}
is overstated; it is only the daily-averaged fields that are to a large extent
in  geostrophic balance (the jPDF of daily-averaged vorticity vs. laplacian of SSH
is a ellipsis with large eccentricity with main axis tilted by 45$^\circ$; figure \ref{fig3}n).

The October vorticity-strain jPDF show much weaker skewness (the vorticity skewness
is 0.68 and 0.67 for hourly and daily-averaged velocities). The shape of the
 vorticity-strain jPDF  appears to be a combination of two half-ellipses centered
 about $\zeta=0$, one with a 45$^\circ$ slope (characteristic of submesocale fronts
 that persist in Summer) and one with a very steep slope. The vorticity-divergence
 jPDF is significantly different than the April one. In particular, the jPDF
 for daily-averaged currents is still egg-shaped but hourly and residual are
 ellipses about $\zeta=0$, likely associated with inertia-gravity waves.
That the submesoscale dynamics in October is mainly unbalanced is clearly depicted
in the shape of jPDF of vorticity-laplacian of SSH for hourly and
residual fields
(figure \ref{fig3}p and \ref{fig3}r), which is an ellipsis aligned in the vertical axis.
Of course daily-averaging the model fields filter out the ageostrophic flows and
the daily-averaged flow is essentially geostrophic.

@@

To better understand the projection of these flows onto different horizontal
scales, we calculate wavenumber spectra of kinetic energy and sea-surface height
(SSH) variance.  Before calculating the spectra, time-mean and spatial linear
trends were removed, and the ensuing fields were multiplied by a Hanning spectral ``window''.
We discuss only spectra for the 1/48$^\circ$, which simply extends the 1/24$^\circ$
towards smaller scales (supplemental material).

Figure \ref{fig3}a depicts the horizontal wavenumber spectra of surface KE in
April and October. At scales larger than 20km, the spectra based on hourly
velocity snapshots are nearly indistinguishable within 95$\%$ confidence level
(compare solid lines in figure \ref{fig3}a). Consistent with the results of
\citet{rocha_etal2016} who analyzed a 5-month output of the LLC4320 simulation in Drake Passage,
there is significant high-frequency variabilty at submesoscales. Daily-averaging
the velocity field suppresses spatial variability at scales smaller than about 250
km both in April and October (compare solid lines against dashed lines in figure
\ref{fig3}a). But this suppression is more dramatic in October, when the horizontally
divergent flows peak. At scales
smaller than 100 km, 39$\%$ of the surface KE in April is accounted for by super-inertial
flows as opposed to 79$\%$ in October. The seasonality of subinertial submesoscale flows
is more dramatic than the seasonal cycle of the total flow. This is consistent
with the results of \citet{sasaki_etal2014}, which are
based on daily-averaged velocity fields of a different model without tidal forcing
(P. Klein, personal communication).

These high-frequency submesoscale flows significantly project on the sea-surface.
In October, when the inertia-gravity waves are strongest the surface and submesoscale
turbulence is weakest, there is a dramatic difference between the spectra
based on hourly and daily-averaged SSH: at scales smaller than 100 km, the spectra
of hourly SSH roughly follows a $-2$ power-law, whereas the spectra of daily-averaged
SSH roughly follows a $-5$ power-law. At these scales, 33$\%$ of the SSH variance in April is accounted for by suprainertial
flows as opposed to 83$\%$ in October! Curiously, the out-of-phase seasonal cycle of submesoscale
turbulence and near-surface inertia-gravity waves conspire to yield very weak seasonality
in the spectra of KE and SSH variance based on hourly field.

\section{Summary and Conclusion}
Our main finding is that
near-surface upper-ocean inertia-gravity waves undergo
a strong seasonal cycle that is out-of-phase with the seasonal cycle of
the submesoscale turbulence. \cite{dasaro1978} showed that linear internal waves
velocity in the mixed-layer strongly depends on the density jump at the mixed-layer
base, with largest mixed-layer velocities when the jump is strongest.
In Summer, the shallow mixed-layer overlays a strong seasonal pycnocline,
and thus the internal waves projection onto the mixed-layer may be stronger
according to \cite{dasaro1978}'s arguments. An
alternative explanation is that the shape baroclinic modes experience significant
seasonal modulation, with significant summertime peak in their surface amplitude
 (supplemental material).
If the internal wave source has weak seasonal dependence, then a strong
seasonal cycle in the near-surface expression of internal waves is expected.
Internal tide generation does not show a seasonal modulation \citep[e.g.,][]{alford2003},
and therefore it is plausible to expect a strong seasonality in the the near-surface
expression of internal tides and
other small-scale internal waves generated through internal tide interactions.
Near-inertial wave generation peaks in Winter, but those waves project on large
horizontal scales \citep[e.g, ][]{qi_etal1995}.

@@

%  this should go in a supplementary material
%\clearpage

%\section{Supplemental material 1: Spectral errors}

\newpage
\section*{Supplemental material 1: On the LLC simulations}

Table \ref{tab:llc} show details of the LLC spin-up hierarchy.
Both simulation analyzed in the present paper, LLC2160 and LLC4320,
were forced by surface fluxes are from the 0.14$^\circ$ European Centre for
Medium-range Weather Forecasting (ECMWF) atmospheric operational model analysis,
 starting 2011 humidity and the 16 most significant  tidal components.
The LLC control files are available online
(at \texttt{http://mitgcm.org/viewvc/MITgcm/MITgcm$\_$contrib/
llc$\_$hires}).

For the LLC hierarchy, the sea-surface height is calculated
using a linearized equation. While the nominal horizontal resolution  of the
LLC4320 simulation is 1/48$^\circ$,
horizontal wavenumber spectra suggests an effective resolution of about 8 km.
(see figure \ref{fig3}), while wavenumber spectra suggests an effective resolution
of 20 km for  LLC2160 simulation.

\begin{table*}[ht]
\label{tab:llc}
\caption{\small The MITgcm Latitute-Longitude Cap (LLC) spin-up hierarchy. This study will analyze
         the output of LLC2160 and LLC4320. Model fields are available hourly.
         Adapted from \cite{rocha_etal2016}.}
\begin{center}
\begin{tabular}{ | c | c | c | c | c |}
\hline
Simulation & Resolution & Time-step & Period  & Tides \\ \hline
ECCO2 Adjoint & $1/6^{^\circ}$ & 1200 s & January 2009 - December 2011 & No\\
LLC 1080   & $1/12^{^\circ}$ & 90 s & January 2010 - July 2012 & No\\
LLC 2160   &  $1/24^{^\circ}$ & 45 s  & January 2011 - April 2013 & Yes\\
LLC 4320   &  $1/48^{^\circ}$ & 25 s & September 2011 - September 2012 & Yes \\
\hline
\end{tabular}
\end{center}
\end{table*}

\subsection*{Spectra}
Kinetic energy and sea-surface height variance wavenumber spectra for the LLC2160
and LLC4320 simulations are shown in figure \ref{figS2_3}. As expected, and
consistent with the submesoscale regional study by \cite{capet_etal2008a}, the
higher-resolution simulation (LLC4320) essentially extends the spectra of the
lower-resolution simulation (LLC2160) towards smaller scale, demonstrating that
the higher resolution is resolving a wider array of scales. Between 20 and 100 km
there is linewidth agreement between the spectra of the two simulations.

\subsection*{Seasonality in statistics: probability density functions}
\begin{figure*}[ht]
   \begin{center}
     \includegraphics[width=1.\textwidth]{figs/fig_s2_2.pdf}
  \caption{Time dependence of probability density functions of vertical vorticity (a, b)
  and horizontal divergence (c, d). Monthly averages in April and October are
  also shown (e, f).}
  \label{figS2_2}
  \end{center}
\end{figure*}

\subsection*{Seasonality in the upper-ocean stratification}

Figure \ref{figS2_1} shows a comparison between potential density sections
at 165$^\circ$E from
the Roemmich-Gilson Argo Climatology \citep[updated from][]{roemmich_gilson2009}.
and the LLC2160 and LLC4320 simulations. Of course, the Argo climatology is smoother
owing to lower resolution and long-term averaging. Nonetheless, simulations reproduce well the strong
seasonality in the stratification. The upper ocean is weakly stratified in late Winter/early Fall
and restratifies in Summer. Argo climatology and LLC simulations mixed-layer depths are consistent
at all seasons (not shown).

@@

\subsection*{How well does the model captures high-frequency modes?}
To assess how well the LLC2160 represents the high-frequency variability,
we compare the model near-surface velocity field against two available
mooring data: KEO (the data is available online at \texttt{http://www.pmel.noaa.gov/ocs/KEO})
and KESS mooring $\#$7 (the data is available online at \texttt{http://uskess.whoi.edu}).
For proper comparison, we use model data closest to these two moorings (west of
the region considered in the present study). Focusing on high-frequencies, we
calculate frequency spectral estimates every month. There are about 48 independent
spectral realizations.

@@

\section*{Supplemental material 3: Seasonality in pressure modes near-surface amplitude}
